TODO Title

### 2013-05-03 CGP Journal/main.tex
Added description of how Reorder works, and pseudo code

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

@@

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
% Do not put math or special symbols in the title.
\title{TODO Title}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

@@

\subsection{Reorder}

@@

@@

\section{Analysis of Evolutionary Mechanisms}
TODO Explain footprint concept here.
\subsection{Inactive Nodes and Node Duplication}
TODO Look at how much of the genome is duplicated in both active and inactive

>>>>

% *** SPECIALIZED LIST PACKAGES ***
%
\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

@@

\usepackage{verbatim}
\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
% Do not put math or special symbols in the title.
\title{TODO Title}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

@@

\subsection{Reorder}
The concept behind \emph{Reorder} is to randomly shuffled node ordering without
impacting gene expression.  This is possible because for a given node, there can
be a large number of other nodes which neither have no required ordering,
as they neither directly or indirectly take input from the node nor provide their
output to the node.  In general, this process works by assigning nodes new
locations in the genome at random once all of the nodes they take input from have been assigned
locations earlier in the genome.

The first step in performing \emph{Reorder} on a genome is to create convenience
data structures to store direct connection relationships, such that given a node
we can determine in constant time where it gets its input and which nodes read
its output.  These structures can be build in order $O(N)$ time, where $N$ is the
number of nodes.

With this preprocessing complete, the algorithm given in Figure~\ref{fig:reorder}
is used to assign each node a new location in the modified genome.
It starts by constructing the \emph{addable} set, which contains all nodes who's direct
dependencies have already been added.  Initially this contains only the input locations.
Once a node is assigned a location, all of the nodes that depend on its output
have that dependency marked as satisfied.  Once all of a nodes dependencies
are satisfied, it can be put into \emph{addable}.  In this way, nodes are
randomly removed from \emph{addable}, assigned the next possible location
in the genome, with new nodes added to \emph{addable} as it becomes semantically
viable.  Iteration ends when all nodes have been assigned new locations.
This algorithm requires $O(AN)$, where $A$ is the arity used by the nodes
and $N$ is the number of nodes.

\begin{figure}
  \begin{algorithmic}

    \STATE $addable \leftarrow input\_locations$
    \STATE $index \leftarrow |input\_locations|$
    \STATE $new\_loc \leftarrow \emptyset$
    \WHILE{$|addable| > 0$}
      \STATE $working \leftarrow randpop(addable)$
      \IF{$working \notin input\_locations$}
        \STATE $new\_loc[working] \leftarrow index$
        \STATE $index \leftarrow index + 1$
      \ENDIF
      \FORALL{$link \in feeds\_to(working)$}
        \STATE Satisfy $link$'s dependence on $working$
        \IF{$|unsatisfied(link)| = 0$}
          \STATE $addable \leftarrow addable \cup \{link\}$
        \ENDIF
      \ENDFOR
      \RETURN $new\_loc$
    \ENDWHILE
  \end{algorithmic}
  \caption{Algorithm used by \emph{Reorder} to determine new node locations.}
  \label{fig:reorder}
\end{figure}

The final step in \emph{Reorder} is to use the list of new locations to convert
the existing genome into the reordered genome.  This can be done in the straight
forward manner by converting connection genes and output location genes using
the $new_loc$ map returned from the algorithm in Figure~\ref{fig:reorder}.

@@

TODO Explain reruning.

@@

\section{Analysis of Evolutionary Mechanisms}
TODO Explain footprint concept here.

\subsection{Inactive Nodes and Node Duplication}
TODO Look at how much of the genome is duplicated in both active and inactive
